To be FAIR: Theory Specification Needs an Update

### 2025-06-19 test.Rmd
update 2025-06-19 16:20:01.419007

Merely publishing theory in a journal article does not make it open;
to be open, theory should adhere to established open science standards for specification and archival.
We propose to implement theories as digital objects,
and archive these with appropriate metadata in a FAIR-compliant repository (e.g., Zenodo).
Metadata are "data about the data". 
They provide information about the nature and content of a digital object and are stored in the repository where the version of record of the FAIR theory is deposited.
FAIR theories are *Findable* via a DOI or by searching the repository they are archived in;
*Accessible* in a machine- and human-readable filetype;
*Interoperable* for specific purposes, for example, within the data analysis environment;
and *Reusable* in the practical and legal sense, so that they may be iteratively improved by the author or by others.
<!-- AB I think some initial word on what is covered is needed here; otherwise readers may wonder what we mean to cover--> 
<!-- CJ: Since this is in the next paragraph, I think it's redundant. Will merge there -->
Following the original proposal of Lamprecht and colleagues [-@lamprechtFAIRPrinciplesResearch2019],
we adapt the FAIR principles for theory, see [Supplemental Table S1](https://github.com/cjvanlissa/fair_theory/blob/main/fair_principles.csv).
We reflect on the necessary (minor) changes,
as well as on the current state and future of FAIR theory in psychology.
The resulting principles provide guidance for instantiating theory as a FAIR digital object,
and we provide worked examples to encourage their adoption.

>>>>

Merely publishing theory in a journal article does not make it open;
to be open, theory should adhere to established open science standards for specification and archival.
We propose to implement theories as digital objects,
and archive these with appropriate metadata in a FAIR-compliant repository (e.g., Zenodo).
Metadata are "data about the data". 
They provide information about the nature and content of a digital object and are stored in the repository where the version of record of the FAIR theory is deposited.
FAIR theories are *Findable* via a DOI or by searching the repository they are archived in;
*Accessible* in a machine- and human-readable filetype;
*Interoperable* for specific purposes, for example, within the data analysis environment;
and *Reusable* in the practical and legal sense, so that they may be iteratively improved by the author or by others.
Following the original proposal of Lamprecht and colleagues [-@lamprechtFAIRPrinciplesResearch2019],
we adapt the FAIR principles for theory, see [Supplemental Table S1](https://github.com/cjvanlissa/fair_theory/blob/main/fair_principles.csv).
We reflect on the necessary (minor) changes,
as well as on the current state and future of FAIR theory in psychology.
The resulting principles provide guidance for instantiating theory as a FAIR digital object,
and we provide worked examples to encourage their adoption.
